Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-27 paper.qmd
Added Figure 5

## To obtain the standardise counts within hexbins
counts_df <- hb_obj_two_curvy$std_cts
df_bin_centroids <- extract_hexbin_centroids(
  centroids_df = all_centroids_df, 
  counts_df = counts_df) |>
  filter(drop_empty == FALSE)

umap_data_two_curvy_with_hb_id <- hb_obj_two_curvy$data_hb_id
df_all_two_curvy <- dplyr::bind_cols(training_data_two_curvy, umap_data_two_curvy_with_hb_id)
df_b_two_curvy <- avg_highd_data(data = df_all_two_curvy, col_start = "x")

@@

```{r}
#| label: empty-bin-two-curvy
hex_grid_nonempty <- hex_grid |>
  filter(hex_poly_id %in% df_bin_centroids$hexID)

hex_grid_nonempty_two_curvy <- ggplot(
  data = hex_grid_nonempty, 
  aes(x = x, y = y)) +
  geom_polygon(color = "black", 
               aes(group = hex_poly_id), 
               fill = "#ffffff") +
  geom_point(data = df_bin_centroids, 
             aes(x = c_x, y = c_y), 
             color = "#33a02c",
             size = 1) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("c", sc_ltr_pos) 
```

<!--2D model-->
```{r}
#| label: triangulate-two-curvy
## Triangulate bin centroids
tr1_object_two_curvy <- tri_bin_centroids(
  df_bin_centroids, x = "c_x", y = "c_y")
tr_from_to_df_two_curvy <- gen_edges(
  tri_object = tr1_object_two_curvy)

@@

@@

```{r}
#| echo: false
#| label: fig-predict-two-curvy
#| fig-pos: H
#| fig-width: 9
#| fig-height: 5
#| fig-cap: "Comparison of prediction generated using the exiting `umap` prediction method and our method: (a) A view of the true model in projections from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, (c) predicted data from our method, and (d) comparison of predictions from both methods. In plot (d), points representing the same data but predicted by different methods are connected by lines. Need to add concluded sentence after changing the data."

# proj_plot_two_curvy_true + plot_predict_umap_two_curvy + 
#   plot_predict_umap_two_curvy + plot_predict_umap_connect_two_curvy +
#   plot_layout(ncol = 4)
proj_plot_two_curvy_true
```

>>>>

## To obtain the standardise counts within hexbins
counts_df <- hb_obj_two_curvy$std_cts
df_bin_centroids_two_curvy <- extract_hexbin_centroids(
  centroids_df = all_centroids_df, 
  counts_df = counts_df) |>
  filter(drop_empty == FALSE)

umap_data_two_curvy_with_hb_id <- hb_obj_two_curvy$data_hb_id
df_all_two_curvy <- dplyr::bind_cols(training_data_two_curvy, umap_data_two_curvy_with_hb_id)
df_bin_two_curvy <- avg_highd_data(data = df_all_two_curvy, col_start = "x")

@@

```{r}
#| label: empty-bin-two-curvy
hex_grid_nonempty <- hex_grid |>
  filter(hex_poly_id %in% df_bin_centroids_two_curvy$hexID)

hex_grid_nonempty_two_curvy <- ggplot(
  data = hex_grid_nonempty, 
  aes(x = x, y = y)) +
  geom_polygon(color = "black", 
               aes(group = hex_poly_id), 
               fill = "#ffffff") +
  geom_point(data = df_bin_centroids_two_curvy, 
             aes(x = c_x, y = c_y), 
             color = "#33a02c",
             size = 1) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("c", sc_ltr_pos) 
```

<!--2D model-->
```{r}
#| label: triangulate-two-curvy
## Triangulate bin centroids
tr1_object_two_curvy <- tri_bin_centroids(
  df_bin_centroids_two_curvy, x = "c_x", y = "c_y")
tr_from_to_df_two_curvy <- gen_edges(
  tri_object = tr1_object_two_curvy)

@@

```{r}
#| label: model-prediction-umap-quollr

true_pred_df <- predict_emb(
  test_data = true_model_df,
  df_bin_centroids = df_bin_centroids_two_curvy,
  df_bin = df_bin_two_curvy,
  type_NLDR = "UMAP"
)

# Compute radius r
a1 <- calc_bins_y(
  bin1 = 15, 
  r2 = r2, 
  q = 0.1)$a1

r <- a1/2

# Function to jitter points within a circumcircle of a hexagon
jitter_within_circumcircle <- function(center_x, center_y, radius, num_points) {
  theta <- runif(num_points, 0, 2 * pi)
  r <- radius * sqrt(runif(num_points))
  x <- center_x + r * cos(theta)
  y <- center_y + r * sin(theta)
  jittered_points <- data.frame(UMAP1 = x, UMAP2 = y, ID = points_in_hex$ID)
  return(jittered_points)
}

# Jittered points data frame
jittered_points_df <- data.frame()

# Iterate through each hexagon and jitter points within the circumcircle
for (hex_id in unique(true_pred_df$pred_hb_id)) {
  # hex_points <- hex_grid_nonempty %>% filter(hex_poly_id == hex_id)
  # center_x <- mean(hex_points$x)
  # center_y <- mean(hex_points$y)
  # points_in_hex <- true_pred_df %>% filter(pred_UMAP_1 >= min(hex_points$x) & pred_UMAP_1 <= max(hex_points$x) & pred_UMAP_2 >= min(hex_points$y) & pred_UMAP_2 <= max(hex_points$y))

  points_in_hex <- true_pred_df |>
    filter(pred_hb_id == hex_id)

  center_x <- points_in_hex |> 
    pull(pred_UMAP_1) |> 
    unique()

  center_y <- points_in_hex |> 
    pull(pred_UMAP_2) |> 
    unique()

  if (nrow(points_in_hex) > 0) {
    jittered_points <- jitter_within_circumcircle(center_x, center_y, r, nrow(points_in_hex))
    jittered_points_df <- rbind(jittered_points_df, jittered_points)
  }
}

plot_predict_umap_two_curvy <- ggplot(
  data = jittered_points_df, 
  aes(x = UMAP1,
      y = UMAP2)) +
  geom_point(alpha = 0.5) +
  interior_annotation("c", c(0.92, 0.96), cex = 2)
```

@@

<!--prediction for true-model of two-curvy clust-->
```{r}
#| label: model-prediction-umap-original

predict_umap_two_curvy <- read_rds(file = "data/two_curvy_clust/two_curvy_clust_umap_predict_true.rds")

predict_two_curvy_obj <- gen_scaled_data(
    data = predict_umap_two_curvy)

predict_umap_two_curvy_scaled <- predict_two_curvy_obj$scaled_nldr

plot_predict_umap_fun_two_curvy <- ggplot(
  data = predict_umap_two_curvy_scaled, 
  aes(x = UMAP1, 
      y = UMAP2)) +
  geom_point(alpha = 0.5) +
  interior_annotation("b", c(0.92, 0.96), cex = 2)
```

```{r}
#| label: model-prediction-points-line

predict_umap_two_curvy_scaled <- predict_umap_two_curvy_scaled |>
  mutate(ID = row_number()) |>
  mutate(type = "predict_from_umap")

jittered_points_df <- jittered_points_df |>
  mutate(type = "predict_from_model")

predict_point_df <- bind_rows(
  predict_umap_two_curvy_scaled,
  jittered_points_df
)

plot_predict_umap_connect_two_curvy <- ggplot(
  data = predict_point_df, 
  aes(x = UMAP1, 
      y = UMAP2, 
      group = ID)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.3) +
  interior_annotation("d", c(0.92, 0.96), cex = 2)
```

```{r}
#| echo: false
#| label: fig-predict-two-curvy
#| fig-pos: H
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Comparison of prediction generated using the exiting `umap` prediction method and our method: (a) A view of the true model in projections from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, (c) predicted data from our method, and (d) comparison of predictions from both methods. In plot (d), points representing the same data but predicted by different methods are connected by lines. Need to add concluded sentence after changing the data."

proj_plot_two_curvy_true + plot_predict_umap_fun_two_curvy +plot_predict_umap_two_curvy +
  plot_predict_umap_connect_two_curvy +
  plot_layout(ncol = 2)
```
